### 2015-08-12 Proposals/NERC-CCUS_2015/Outline_ReactiveCCS.tex
Updates in the proposal

%%%
%%% space between lines
%%%
\renewcommand{\baselinestretch}{1.}

@@

@@

\abstract{In Carbon Capture, Utilisation, and Storage (CCUS) integrated processes, \CO\; is captured from energy-intensive point sources, mitigated by either chemical conversion or injection into geological formations. Each CCUS process is at very different stages of development and optimisation. In this project, research into developing an integrated dynamic prediction tool model will yield a capability to monitor migration of \CO\; plumes in geological storage sites. The generic mathematical model (High-order accurate Control Volume Finite Element Methods / HPC Computational Multi-Fluid Dynamics) will enable detailed investigation of: (a) \CO\; transport in heterogeneous media, (b) \CO-brine-rock interactions (dissolution, precipitation and exsolution due to thermal gradients and brine dissolution in supercritical \CO) and impact on permeability / porosity spatial- and time-variability.  The integrated model will be validated against

 We have assembled a team which we believe can successfully carry out the proposed work, combining a wide range of skills encompassing physics, mathematics, earth science and chemical engineering.
}

@@

>>>>

%%%
%%% space between lines
%%%
\renewcommand{\baselinestretch}{1.3}

@@

\abstract{Environmental and industrial porous media flow is key to a wide range of traditional and emerging applications, including hydrocarbon exploration, geothermal reservoir engineering, nuclear waste repository, groundwater contamination, protection and remediation and \CO\; capture and storage. In this project, research into developing an integrated dynamic prediction tool model will yield a capability to monitor migration of \CO\; plumes in geological storage sites. The generic mathematical model (High-order accurate Control Volume Finite Element Methods / Computational Multi-Fluid Dynamics) will enable detailed investigation of: (a) \CO\; transport in heterogeneous media, (b) \CO-brine-rock interactions (dissolution, precipitation and exsolution due to thermal gradients and brine dissolution in supercritical \CO) and impact on permeability / porosity spatial- and time-variability.  The integrated model will be validated against (a) standard SPE reservoir simulation benchmarks and (b) field data obtained from existing experiments (e.g., Frio Brine Pilot Experiment and XXXX).  We have assembled a team which we believe can successfully carry out the proposed work, combining a wide range of skills encompassing physics, mathematics, earth science and engineering. }

@@

@@

\begin{comment}
This proposal comes at a time when the UK nuclear sector is resurgent. It is now widely accepted that one of the ways in which the UK can meet its commitments to reducing CO2 emissions, as well as dealing with its over-reliance on imported fuel supplies, is to replace the current fleet of ageing nuclear reactors.

Natural and industrial porous media is key to a wide variety of traditional and emerging engineering applications, including but not limited to oil and gas extraction from geological reservoirs, carbon capture and storage, geothermal reservoir engineering, soil sciences, groundwater remediation and protection, biological engineering, food processing, fuel cells, nano-technology, construction engineering, wood processing and printing.

Global warming due to the accumulating concentration of atmospheric CO2 is being accepted by the scientific and broader community. Amongst a number of CO2 atmospheric emission management considerations is carbon dioxide capture and storage, CCS, related to major coal burning power generation. The operation of CCS requires a close synergy of CO2 capture and CO2 storage. Clearly the most effective capture technology is of no value unless the captured CO2 can be stored. Carbon dioxide is a very soluble gas which makes storing CO2 in large volumes of liquids in porous subsurface reservoirs a serious consideration for captured CO2. Subsurface brine aquifers have a large potential for CO2 storage. Computer based reservoir simulation has been used increasingly by oil companies to predict flow and oil recovery behaviour of reservoirs and help in decision making of project viability. The quality of the simulation process depends on the ability of the simulator to represent the physics and chemistry of the various phenomena occurring in the reservoir and the quality of the various properties characterising the reservoir structure and fluids. The Institute of Petroleum Engineering at Heriot-Watt University will use its oil related reservoir simulation experience and existing commercial simulation software to predict the behaviour of injecting carbon dioxide into brine containing structures in the near vicinity of power generation stations. Predicting the behaviour of CO2 injected into subsurface rocks containing brine requires an understanding of the impact of a wide range of issues; the geology of the formation, the properties of the porous rocks and the fluids, and the change in these properties resulting from interaction with CO2. Reservoir simulation provides an effective platform to evaluate these issues and generate forecasts of CO2 storage potential and distribution. Integrating with and using data from BGS and UoE, the Heriot-Watt team will construct reservoir simulation models, using different levels of characterisation, detail, to simulate specific potential storage sites. The team will examine the impact of a wide range of characterising parameters. Injecting CO2 into brine containing rocks is likely to change rock properties affecting flow behaviour. For example the dissolution of acidic CO2 saturated brine could change the mechanical strength of the rock as well as dissolve minerals, again impacting on flow behaviour. As well as predicting this impact in the reservoir simulation process, the Heriot-Watt team will carry out laboratory based evaluations on rocks, representative of the potential sites, on mechanical strength changes as a result of CO2 brine interaction. Using reservoir condition experimental facilities evaluation of chemical interactions using static soaking and dynamic flow of fluids through rocks (core flooding) and measuring effluent ion concentrations and permeability changes, the HWU team will examine the impact of chemical interaction of carbonated brine with the various rock types provided by BGS. The output from these reservoir simulation based evaluations will be integrated with the other partner outputs and use will be made of 3D visualisation to enable project partners to examine the impact of various issues on the short and long term storage capacity and distribution of the considered subsurface storage sites. The output and 3D simulation will also be used to communicate to other interested stake holders, eg. Scottish Government, DTI, Local Government, EC, other power generators, environmental agencies, and not least the general public, the process and potential of subsurface CO2 storage.
